Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-18 grove-paper.tex
add incomplete decomposition algorithm

%% Packages
\usepackage{booktabs}
\usepackage[rule=false]{subcaption}
\usepackage{graphicx}
\usepackage{semantic}
\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
\let\colonsim\undefined % Avoid redefinition error in `colonequals`
\usepackage{colonequals}
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{fontawesome}
\usepackage{array}
\usepackage{todonotes}

@@

% math ligatures
\mathlig{::=}{\coloncolonequals}
\mathlig{|-}{\vdash}
\mathlig{:->}{\mapsto}
\mathlig{-->a}{\overset{a}{\longrightarrow}}
\newcommand{\abs}[1]{\left|{#1}\right|}

@@

@@

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid * \\
      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda q \mathord{:}\tau.e \mid e~e \mid e~op~e \mid n \\
                                  &      &           &         & \phantom{x} \mid \hole \mid \{e|\ldots|e\} \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole \mid \{ \tau | \ldots | \tau \} \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  \label{fig:Syntax as a grammar}
\end{figure}

@@

@@

If we translate the graph into expressions in the calculus, interactions with
the graph can be modeled as a stream of user actions as described in
\autoref{sub:User Actions}. Our decomposition algorithm partitions the graph
into disjount connected components. We encode cycles by resolving all but one
vertex of each multi-edge into a by-id reference to the one vertex.

>>>>

%% Packages
\usepackage{booktabs}
\usepackage[rule=false]{subcaption}
\usepackage{graphicx}
\usepackage{semantic}
\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
\let\colonsim\undefined % Avoid redefinition error in `colonequals`
\usepackage{colonequals}
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{fontawesome}
\usepackage{array}
\usepackage{todonotes}
\usepackage[ruled]{algorithm2e}

@@

% math ligatures
\mathlig{::=}{\coloncolonequals}
\mathlig{:=}{\colonequals}
\mathlig{|-}{\vdash}
\mathlig{:->}{\mapsto}
\mathlig{-->a}{\overset{a}{\longrightarrow}}
\newcommand{\abs}[1]{\left|{#1}\right|}

@@

% TODO replace screenshots with text
% TODO make ref code gray
% TODO move whole figure left
% TODO remove cursor from screenshots
% TODO add multiparent box
% TODO only show parens where needed - figure 10, 11 - by associativity
% TODO repplace cycle screenshot with cycles box 

@@

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Operators:}         & op   & {}\in Ops & {}::={} & + \mid * \\
      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x \mid \hole \mid \{q|\ldots|q\} \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda q \mathord{:}\tau.e \mid e~e \mid e~op~e \mid n \\
                                  &      &           &         & \phantom{x} \mid \hole \mid \{e|\ldots|e\} \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole \mid \{ \tau | \ldots | \tau \} \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  % TODO add references into multiparent store and cycles store
  \label{fig:Syntax as a grammar}
\end{figure}

@@

% NOTE base grammar has holes. rooted expression grammar has refs, conflicts, and multiparents
% NOTE ref = expression reference or vertex reference

@@

Translating a graph into expressions in the calculus allows us to model
user interaction as a stream of edits as described in \autoref{sub:User
Actions}. Our decomposition algorithm partitions the graph into disjount
connected components. We encode cycles by resolving all but one vertex of each
multi-edge into a by-id reference to the one vertex.

\begin{algorithm}[t]
  \caption{Graph Decomposition}
  \label{fig:Graph Decomposition}

  \DontPrintSemicolon
  \SetAlgoNoLine
  \SetKwFunction{reachable}{reachable}
  \SetKwFunction{simpleCycles}{simpleCycles}
  \SetKwProg{Fn}{function}{}{end}

  \KwIn{Graph $G$}
  \KwOut{Decomposition of $G$ into cycle-free expressions}

  $liveEdges := \{\e : G(\e) = \Plus\}$\;
  $liveVertexes := \bigcup_{(u, v, p, v') \in liveEdges} \{v, v'\}$\;
  $allVertexes := \bigcup_{(u, v, p, v') \in \dom(G)} \{v, v'\}$\;
  $multiparented := \{v \in allVertexes : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in \dom(G), v_1 \ne v_2\}$\;
  $deleted := allVertexes \setminus liveVertexes \setminus \{v_0\}$\;
  $MP := \{\reachable(v, \varnothing) : v \in multiparented\}$\;
  $D := \{\reachable(v, \varnothing) : v \in deleted\}$\;
  \Return{$(\reachable(v_0, \varnothing), MP, D, \simpleCycles(deleted))$}\;
  \;

  \Fn{\reachable $((u, k), seen)$}{
    \If{$(u, k) \in multiparented \cup seen$}{
      \Return{$u$}
    }
    $childEdges := \{(u', v, p, v') \in liveEdges : v = (u, k)\}$\;
    $children := \{(p, \reachable(v, seen \cup \{u\})) : (p, v) \in childEdges\}$\;
    \Return{$(k, children)$}
  }
  \;

  \Fn{\simpleCycles $(remaining)$}{
    \Return{}
  }
  \;
\end{algorithm}
